---
title: "text_compiler_&_bigrams_TLA"
author: "Benji Leung-Arellano"
date: "2026-09-30"
output: html_document
---

# Loading libraries
```{r}
library(tidyverse)
library(janitor)
library(striprtf)
library(tidytext)
library(rio)
library(dplyr)

```

######################
### TEXT COMPILING ###
######################

# Reformat .RTF files

```{r}
# Load required packages

# Set the paths for your folders
input_folder <- "./Deaf-Technology-Sentiment-NexisUni-Database/Articles/"
output_folder <- "./DTS_articles_raw_text/" # Replace with your output folder path

# Create output folder if it doesn't exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}

# Get a list of all .rtf files in the input folder
rtf_files <- list.files(path = input_folder, pattern = "\\.RTF$", full.names = TRUE)

# Convert each .rtf file to .txt
for (file in rtf_files) {
  # Extract the file name without extension
  file_name <- tools::file_path_sans_ext(basename(file))
  
  # Read the RTF content
  rtf_content <- read_rtf(file)
  
  # Create output file path
  output_file <- file.path(output_folder, paste0(file_name, ".txt"))
  
  # Write the content to a .txt file
  writeLines(rtf_content, output_file)
  
  # Print progress
  cat("Converted:", file, "to", output_file, "\n")
}

cat("Conversion complete!\n")
```


# Raw text compiler

```{r include=FALSE}
#This creates an index with the file path to the stories. And then it compiles the stories into a dataframe
#####################
# Begin SM Code #####
#####################

###
# List out text files that match pattern .txt, create DF
###

#Adjust this line for your file name
files <- list.files("./DTS_articles_raw_text", pattern="*.txt") %>% 
  as.data.frame() |> 
  rename(filename = 1) |> 
  #create an matching file name
  mutate(index = str_replace_all(filename, ".txt", "")) %>%
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", index))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25)) |> 
  distinct(index, .keep_all = TRUE)


#Join the file list to the index

final_data <- rio::import("Excel_Deaf_AND_Technology_AND_Sentiment.xlsx") |> 
  clean_names() |> 
   #create an matching file name
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", title))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25)) |> 
  distinct(index, .keep_all = TRUE)
```

### Check for duplicate entries

```{r}

final_data |> 
  count(title) |> 
  arrange(desc(n))
```


```{r}
dupe_data <- rio::import("Excel_Deaf_AND_Technology_AND_Sentiment.xlsx") |> 
  clean_names() |> 
   #create an matching file name
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", title))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25))
```
#Counting duplicates
```{r}
dupe_data |> 
  count(title) |> 
  arrange(desc(n)) |> 
  filter(n >1)

```
#Total duplicates
```{r}
dupe_data |> 
  count(title) |> 
  arrange(desc(n)) |> 
  filter(n >1) |> 
  summarize(sum(n))
```

#Create final index
```{r}

final_index <- final_data |> 
  inner_join(files, c("index")) |> 
#you need the actual hard-coded path on this line below to the text
  mutate(filepath = paste0("./DTS_articles_raw_text/", filename))

head(final_index)
```

#Fact Check

```{r}

anti_final_index <- final_data |> 
  anti_join(files, c("index"))

```

#Checking for duplicates

```{r}
final_index |> 
  count(title) |> 
  arrange(desc(n))

```

#Text compiler

```{r}
###
# Define function to loop through each text file 
###

create_article_text <- function(row_value) {
  
  #row_value is the single argument that is passed to the function
  # Take each row of the dataframe
  temp <- final_index %>%
    slice(row_value)
  
  # Store the filename for  use in constructing articles dataframe
  temp_filename <- temp$filename
  
  # Create a dataframe by reading in lines of a given textfile
  # Add a filename column 
  articles_df_temp <- read_lines(temp$filepath) %>%
    as_tibble() %>%
    mutate(filename = temp_filename)
  
  # Bind results to master articles_df
  # <<- returns to global environment
  articles_df <<- articles_df %>%
    bind_rows(articles_df_temp)
}

###
# Create elements needed to run function
###

# Create empty tibble to store results
articles_df <- tibble()
#running once to test
#create_article_text(2) 
# Create an array of numbers to loop through, from 1 to the number of rows in our index dataframe 
row_values <- 1:nrow(final_index)

###
# Execute function using lapply
# This loops through each row of the dataframe and append results to master file
###

lapply(row_values, create_article_text)

###
# Clean up articles_df and join to index dataframe
###

articles_df <- articles_df %>%
  select(filename, sentence=value) %>%
  inner_join(final_index)


write_csv(articles_df, "DTS_articles.csv")

```



###############
### BIGRAMS ###
###############

# We'll reuse articles_df even though we've imported it as a csv -> we can skip importing

# Tokenizing sentences of articles, sentence column
```{r}
bigrams <- articles_df |> 
  unnest_tokens(bigram, sentence, token="ngrams", n=2)

bigrams
```

#Remove stopwords and count the words
```{r}

data(stop_words)

bigrams |> 
  filter(!is.na(bigram)) |> 
  separate(bigram, c("word1", "word2"), sep = " ") |> 
  filter(!word1 %in% stop_words$word) |> 
  filter(!word2 %in% stop_words$word) |> 
  count(word1, word2)

```

# Create Bigrams

```{r}
#Filter out stop words.

bigram_filtered <- bigrams |> 
  filter(!is.na(bigram)) |> 
  separate(bigram, c("word1", "word2"), sep = " ") |> 
  filter(!word1 %in% stop_words$word) |> 
  filter(!word2 %in% stop_words$word) |> 
  count(word1, word2, sort = TRUE)
```

```{r}
bigram_filtered
```

# Saving dataframes to the folder
```{r}
write_csv(bigram_filtered, "bigram_filtered.csv")
```

```{r}
deaf_bigram <- bigram_filtered |> 
  filter(grepl("deaf", word1, ignore.case = TRUE) | grepl("deaf", word2, ignore.case = TRUE))
```


############
### MEMO ###
############

My bigram findings indicate a large focus in the articles on artificial intelligence. The term “artificial intelligence” has the highest count in the bigram table, with 320 entries. The runner-up for that position is “venture capital”, which is not related to the topic I am interested in. Upon scrolling through the bigram table more, it was found that the contents of these bigrams primarily focus on finance topics rather than my intended goal of topics surrounding Deaf sentiments on technology (DTS). To face this problem, I wrote a piece of code to filter through word1 and word2 columns in the bigram to find any bigrams with the keyword deaf in them, so I could get any wildcard variation of deaf while trying to obtain some findings. Out of 12706 rows, it was found that only 46 rows had the keyword deaf in them. I didn’t expect many articles to focus on that topic to begin with, both bigrams show that the dataset reflects that expectation. I am aware that my original search was faulty for the goal I was searching for, if I use a better search query I hope I can find sources that focus on DTS rather than sources that briefly mention DTS, such as these financial sources. In the spirit of using the fullest potential of this dataset and considering that the current bigram focuses mostly on financial topics, in the future, when filtering this dataset for purposes similar to the bigram table, I would like to first filter the dataset for sentences that contain the word deaf in them so the queries can focus on deafness, which is the primary component of DTS. 
