---
title: "Basic R Proficiency"
author: "Tonatiuh Leung-Arellano"
date: "2026-09-23"
output: html_document
---

Homework: Basic R Proficiency

# Loading libraries
```{r}
library(tidyverse)
library(janitor)
```


# Loading in mob analysis data & cleaning column names
```{r}
mob_data <- rio::import("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/mob_analysis_data.csv")

mob_data <- clean_names(mob_data)
  
```

# Number of rows and columns
```{r}
nrow(mob_data)
ncol(mob_data)
```

# Counting the top 15 newspaper entries
# slice_max allows me to do top 15, with_ties removes any ties at the end (there were 17 entries with ties allowed, we want 15)
```{r}
mob_data |> 
  count(newspaper) |> 
  arrange(desc(n)) |> 
  slice_max(order_by = n, n = 15, with_ties = FALSE)
```



# Counting different entires in 'keyword' column
```{r}
mob_data |> 
  count(keyword) |> 
  arrange(desc(n))
```

# Counting specific articles by year
```{r}

```

# ggplot for total entries by year
```{r}

```

